---
title: "Analyzing the Iris dataset"
author: "Orlando Karam"
date: "Thursday, October 09, 2014"
output: html_document
---

The iris dataset is a dataset of the length and width of petals and sepals (the kind of leaves under the petals) for 3 different varieties of irises. It has been used as a standard dataset to compare ML and statistics techniques for a long time.

It is included in R, in the datasets package. We can usually access it as iris.

The dataset has 150 observations (rows), 50 for each variety, and has the following columns: Petal.Width, Petal.Length, Sepal.Width, Sepal.Length, Species.

```{r}
head(iris)
```

We can plot Sepal width vs height, with different colors for each variety

```{r}
plot(iris$Sepal.Length, iris$Sepal.Width, pch=21, bg=c("red","green","blue")[unclass(iris$Species)], main ="Sepal width vs height");
```
pch 21 tells plot to use colored circles, and bg specifies the colors.

unclass maps the different values to numbers, as you can see below 
```{r}
unclass(iris$Species)
```
Similarly, we can plot petal width and height

```{r}
plot(iris$Petal.Length, iris$Petal.Width, pch=21, bg=c("red","green","blue")[unclass(iris$Species)], main ="Petal width vs height");
```

We can see that setosa is quite easy to distinguish, but virginica and versicolor are more intermixed, especially in their sepal dimensions.


